---
title: "Density Curve Vignette"
author: "Johnny Rodriguez"
date: "2022-10-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  warning = FALSE, message = FALSE)
```

### Introduction - Density Curve 
This vignette will demonstrate how to build a density curve given a 538 dataset.  We will plot two factors in the dataset and place them side by side for visual comparison.

The density plot curve visually displays where the area of the curve falls for a distribution.

### Load the Libraries

```{r}
library(tidyverse)
library(ggplot2)
library(gridExtra)
```

### Read the 538 CSV - New Voter Registrations by Jurisdiction

This dataset contains new voter registrations for the 2016 and 2020 elections

```{r}

mydata <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/voter-registration/new-voter-registrations.csv", na.strings=c("","NA"))

```

### Step 1 - Filter the data for the data needed for analysis
The data will be filtered for new voter registrations in 2016.  This data can also be filtered by jurisdiction and month.


```{r}

#Filters for new voter registrations in 2016
new2016voters <- mydata %>% 
  filter(Year == "2016")

head(new2016voters)

```

### Step 2 - Calculate mean and standard deviation
Calculate the data mean and standard deviation using sd and mean functions

```{r}

#Calculate mean
mean2016newvoters <- mean(new2016voters$New.registered.voters) %>% 
  print()

#Calculate the standard deviation
sd2016newvoters <- sd(new2016voters$New.registered.voters) %>% 
  print()

```

### Step 3 - Create the density plot
Given the mean and standard deviation of the distribution, create the density plot with ggplot.  The plot is further enhanced by marking the means and standard deviation values.

```{r}

#Create the histogram and density curve
new2016 <- ggplot(new2016voters, aes(x = New.registered.voters)) +
  geom_blank() +
  
  #Creates the histogram for this distribution
  geom_histogram(aes(y = ..density..), color = "#000000", fill = "#0099F8", binwidth = 10000) +
  
  #Calculates the density curve for the distribution
  stat_function(fun = dnorm, args = c(mean = mean2016newvoters,sd = sd2016newvoters), col = "tomato")+
  
  #Add labels to part of the density plot
  ggtitle("New Registered Voters in 2016")+
  labs(y = "Density", x = "2016 New Registered Voters")+
  geom_vline(aes(xintercept = mean2016newvoters), color = "#000000", size = 1) +
  geom_vline(aes(xintercept = mean2016newvoters + sd2016newvoters), color = "#000000", size = .75, linetype = "dashed") +
  geom_vline(aes(xintercept = mean2016newvoters - sd2016newvoters), color = "#000000", size = .75, linetype = "dashed")
  
new2016

```

### Step 4 - Repeat the process for the comparison plot
We will repeat the same procedure to generate the density plot for New Registered Voters in 2020.  All calculations are performed in one chunk.

```{r}

#Filters for new voter registrations in 2020
new2020voters <- mydata %>% 
  filter(Year == "2020") 

#Calculate mean
mean2020newvoters <- mean(new2020voters$New.registered.voters) 

#Calculate the standard deviation
sd2020newvoters <- sd(new2020voters$New.registered.voters) 

#Create the histogram and density curve
new2020 <- ggplot(new2020voters, aes(x = New.registered.voters)) +
  geom_blank() +
  
  #Creates the histogram for this distribution
  geom_histogram(aes(y = ..density..), color = "#000000", fill = "#0099F8", binwidth = 10000) +
  
  #Calculates the density curve for the distribution
  stat_function(fun = dnorm, args = c(mean = mean2020newvoters,sd = sd2020newvoters), col = "tomato")+
  
  #Add labels to part of the density plot
  ggtitle("New Registered Voters in 2020")+
  labs(y = "Density", x = "2020 New Registered Voters")+
  geom_vline(aes(xintercept = mean2020newvoters), color = "#000000", size = 1) +
  geom_vline(aes(xintercept = mean2020newvoters + sd2020newvoters), color = "#000000", size = .75, linetype = "dashed") +
  geom_vline(aes(xintercept = mean2020newvoters - sd2020newvoters), color = "#000000", size = .75, linetype = "dashed")
  
new2020

```

### Step 5 - Use Grid Extra Package to place them side by side.
This allows for a visual inspection of both election periods.

```{r}

#Side by side grids
grid.arrange(new2016, new2020, ncol = 2)


```



